---
title: "Tooth Growth Basic Inferential Analysis"
author: "Marcus Vinicius Monteiro de Souza"
date: "June 19, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

This work explores the ToothGrowth dataset from the Base R datasets package to find relationships, if there are any, between tooth growth and vitamin C consumption. Specifically, we'll try to answer the following questions:

1. Is there a difference between ascorbic acid and orange juice supplements concerning tooth growth?
2. Is tooth growth affected by vitamin C dosage?

## A first look at the data

The ToothGrowth dataset description reads:

> The response is the length of odontoblasts (cells responsible for tooth growth) in 60 guinea pigs. Each animal received one of three dose levels of vitamin C (0.5, 1, and 2 mg/day) by one of two delivery methods, (orange juice or ascorbic acid (a form of vitamin C and coded as VC).

Let's load the required packages:

```{r load_required_packages}
if (!require('pacman')) {
  install.packages('pacman')
}
pacman::p_load(tidyverse)
```

and the data:

```{r load_data}
data('ToothGrowth')

ToothGrowth <- as_tibble(ToothGrowth) %>%
  mutate(dose = as.factor(dose))

```

Now, let's have a look at it's first observations:

```{r head}
head(ToothGrowth)
```

and it's last ones:

```{r tail}
tail(ToothGrowth)
```

See a compact display of it's contents:

```{r str}
str(ToothGrowth)
```

and a summary of some of it's statistics:

```{r summary}
summary(ToothGrowth)
```

To clarify the meaning of the variables:

Variable Name | Description
--------------|-------------
len           | Tooth length
supp          | Supplement type: VC (ascorbic acid) or OJ (orange juice)
dose          | Dose in milligrams/day

## Is there a difference between ascorbic acid and orange juice supplements concerning tooth growth?

```{r}
ToothGrowth %>%
  group_by(supp, dose) %>%
  summarise(mean.len = mean(len, na.rm = TRUE)) %>%
  ggplot(aes(dose, mean.len)) +
  facet_grid(. ~ supp) +
  geom_point()
```

