---
title: "Tooth Growth Basic Inferential Analysis"
author: "Marcus Vinicius Monteiro de Souza"
date: "June 19, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

This work explores the ToothGrowth dataset from the Base R datasets package to find relationships, if there are any, between tooth growth and vitamin C consumption. Specifically, we'll try to answer the following questions:

1. Is there a difference between ascorbic acid and orange juice supplements concerning tooth growth?
2. Is tooth growth affected by vitamin C dosage?

## A first look at the data

The ToothGrowth dataset description reads:

> The response is the length of odontoblasts (cells responsible for tooth growth) in 60 guinea pigs. Each animal received one of three dose levels of vitamin C (0.5, 1, and 2 mg/day) by one of two delivery methods, (orange juice or ascorbic acid (a form of vitamin C and coded as VC).

Let's load the required packages:

```{r load_required_packages}
if (!require('pacman')) {
  install.packages('pacman')
}
pacman::p_load(tidyverse)
```

and the data:

```{r load_data}
data('ToothGrowth')

ToothGrowth <- as_tibble(ToothGrowth) %>%
  mutate(dose = as.factor(dose))

```

Now, let's have a look at it's first observations:

```{r head}
head(ToothGrowth)
```

and it's last ones:

```{r tail}
tail(ToothGrowth)
```

See a compact display of it's contents:

```{r str}
str(ToothGrowth)
```

and a summary of some of it's statistics:

```{r summary}
summary(ToothGrowth)
```

To clarify the meaning of the variables:

Variable Name | Description
--------------|-------------
len           | Tooth length
supp          | Supplement type: VC (ascorbic acid) or OJ (orange juice)
dose          | Dose in milligrams/day

## Is there a difference between ascorbic acid and orange juice supplements concerning tooth growth?

If there was no difference between ascorbic acid and orange juice supplements concerning tooth growth, we would expect that the average tooth length would be the same, plus or minus chance variation. Let's have another look at the data: 

```{r mean_tooth_len_by_dose_and_supp}
ToothGrowth %>%
  group_by(supp, dose) %>%
  summarise(mean.len = mean(len)) %>%
  ggplot(aes(dose, mean.len)) +
  geom_col(aes(fill = dose)) +
  facet_grid(. ~ supp) +
  ggtitle('Mean Tooth Length by Dose and Supplement Type')
```

It shows that the mean tooth length is higher for orange juice at doses 0.5 and 1, and equal at dose 2. But is that difference [significant](https://en.wikipedia.org/wiki/Statistical_significance)? Or it can be better explained by chance?

### Hypothesis Test: Orange Juice Supplements Lead to More Tooth Growth Than Ascorbic Acid Ones

To compare the two means statistically, we can perform a [t-test](http://www.investopedia.com/terms/t/t-test.asp), providing the following assumptions about data hold true:

1. Normality.

First, let's check if the sample can be considered as drawn from a normal distribution:

```{r is_data_normal}
qqnorm(ToothGrowth$len, main = 'Tooth Length Normal Q-Q Plot')
qqline(ToothGrowth$len, col = 2)
```

The majority of the sample's observations fit the theoretical normal line.

2. Independence

The observations were drawn indep

```{r t_test_len_supp}

```
